R Notebook

### 2021-05-10 FEM_R/calculate_dflys_from_summary.Rmd
Found a better way to calculate disability free and disease free life years from summary

# Read in Data
```{r}
## Read in all the data
# Baselines
base <- read_dta('ELSA_core_base/ELSA_core_base_summary.dta')
cohort <- read_dta('ELSA_core_cohort/ELSA_core_cohort_summary.dta')

# Hearte (cohort and pop)
#heart_c <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_summary.dta')
heart_p <- read_dta('ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')

@@

# BMI
nomorbid <- read_dta('ELSA_core_nomorbid/ELSA_core_nomorbid_summary.dta')
obese2 <- read_dta('ELSA_core_noobese2/ELSA_core_noobese2_summary.dta')
obese1 <- read_dta('ELSA_core_noobese1/ELSA_core_noobese1_summary.dta')

# Drinking
nodrink <- read_dta('ELSA_core_remove_drink_p/ELSA_core_remove_drink_p_summary.dta')
noBadDrink <- read_dta('ELSA_core_remove_problem_drink_p/ELSA_core_remove_problem_drink_p_summary.dta')
```

# Disability Free Life Years

@@

## Try again, this time with the correct variable (not disabled)

>>>>

# Read in Data
```{r}
## Read in all the data
# Baselines
#base <- read_dta('ELSA_core_base/ELSA_core_base_summary.dta')
cohort <- read_dta('ELSA_core_cohort/ELSA_core_cohort_summary.dta')

# Hearte (cohort and pop)
heart_c <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_summary.dta')
#heart_p <- read_dta('ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')

@@

# BMI (cohort)
#nomorbid <- read_dta('ELSA_core_nomorbid/ELSA_core_nomorbid_summary.dta')
#obese2 <- read_dta('ELSA_core_noobese2/ELSA_core_noobese2_summary.dta')
obese1 <- read_dta('ELSA_core_noobese1/ELSA_core_noobese1_summary.dta')

# Drinking (pop)
nodrink_p <- read_dta('ELSA_core_remove_drink_p/ELSA_core_remove_drink_p_summary.dta')
nodrink_c <- read_dta('ELSA_core_remove_drink_c/ELSA_core_remove_drink_c_summary.dta')
#noBadDrink <- read_dta('ELSA_core_remove_problem_drink_p/ELSA_core_remove_problem_drink_p_summary.dta')
```

# IDEA!!!

Think we need to get the average of disabled var and multiply by the total endpop of corresponding year.
This way, we'll calculate the average number of diseased (or disabled) years, and then we can do the 
opposite calculation to get non-diseased or non-disabled.

Got this idea from looking at the `summary_file_statistics.do` file in input_data/detailed_output/, lines 72-77. Check out the whole idea, this script goes:

- Calculate total_lifeyear for each person m_endpop_xxx * 2
- Calculate summed_total_lifeyear for whole population ( sum(total_lifeyear) )
- Calculate average_lifeyear (summed_total_lifeyear / m_endpop_xxx)

Same can be done for disability free life years and disease free life years, see script for details.

## LifeYears

### Functions
```{r}
cohortLifeyear <- function(c) {
  c.tot.lfyrs <- c$m_endpop_all * 2
  c.sum.tot.lfyrs <- sum(c.tot.lfyrs)
  c.avg.lfyr <- c.sum.tot.lfyrs / c$m_endpop_all[1]
  return(c.avg.lfyr)
}

lifeyearComparison <- function(c, i, int.name) {

  c.avg.lfyr <- cohortLifeyear(c)

  i.avg.lfyr <- cohortLifeyear(i)

  print(int.name)
  print(paste('Cohort Average Lifeyears:        ', signif(c.avg.lfyr, digits=4)))
  print(paste('Intervention Average Lifeyears:  ', signif(i.avg.lfyr, digits=4)))
  print(paste('Increase:                        ', signif(x = (i.avg.lfyr - c.avg.lfyr), digits=4)))
}

diseaseVarLifeyear <- function(c, var) {

  prevVar <- paste0('p_', var, '_all')

  c.tot.var.lfyrs <- c[prevVar] * c$m_endpop_all * 2
  c.sum.var.lfyrs <- sum(c.tot.var.lfyrs)
  c.avg.var.lfyrs <- c.sum.var.lfyrs / c$m_endpop_all[1]

  print(paste0(var, ':    ', signif(c.avg.var.lfyrs, digits=4)))
}

### Cohort
```{r}
# Cohort life years
c <- cohort
c.ly <- cohortLifeyear(c)
print(paste0('Cohort Life Years: ', signif(c.ly, digits=4)))
print('-------------------')

# Cohort disease life years
diseaseVarLifeyear(c, 'cancre')
diseaseVarLifeyear(c, 'diabe')
diseaseVarLifeyear(c, 'hearte')
diseaseVarLifeyear(c, 'lunge')
diseaseVarLifeyear(c, 'hibpe')
diseaseVarLifeyear(c, 'stroke')
diseaseVarLifeyear(c, 'alzhe')
diseaseVarLifeyear(c, 'demene')
diseaseVarLifeyear(c, 'smoken')
diseaseVarLifeyear(c, 'drink')
diseaseVarLifeyear(c, 'heavy_smoker')
diseaseVarLifeyear(c, 'problem_drinker')
```

### Smoking
```{r}
# Cohort
c <- cohort
# No smoking
s <- smok

lifeyearComparison(c, s, 'No Smoking')
```

### Heart Disease
```{r}
# Cohort
c <- cohort
# No Heart Disease
h <- heart_c

lifeyearComparison(c, h, 'No Heart Disease')
```

### Obesity
```{r}
c <- cohort
o <- obese1
lifeyearComparison(c, o, 'No Obesity')
```

### Drink
```{r}
c <- cohort
d <- nodrink_c
lifeyearComparison(c, d, 'No Drink')
```

## Disability

### Functions
```{r}
disabilityLifeYearsCohort <- function(c) {
  total_dis <- c$a_disabled_all * c$m_endpop_all
  sum_tot_dis <- sum(total_dis)
  avg_dis <- sum_tot_dis / c$m_endpop_all[1]
  return(avg_dis)
}

disabilityLifeYearComp <- function(c, i, int.name) {

  c.avg.dis <- disabilityLifeYearsCohort(c)

  i.avg.dis <- disabilityLifeYearsCohort(i)

  print(int.name)
  print(paste('Cohort Average Lifeyears:        ', signif(c.avg.dis, digits=4)))
  print(paste('Intervention Average Lifeyears:  ', signif(i.avg.dis, digits=4)))
  print(paste('Increase:                        ', signif(x = (i.avg.dis - c.avg.dis), digits=4)))
}

disabilityFreeLifeYearsCohort <- function(c) {
  total_disfree <- c$a_not_disabled_all * c$m_endpop_all
  sum_tot_disfree <- sum(total_disfree)
  avg_disfree <- sum_tot_disfree / c$m_endpop_all[1]
  return(avg_disfree)
}
```

### Cohort
```{r}
# Cohort life years
c <- cohort
c.dly <- disabilityLifeYearsCohort(c)
print(paste0('Cohort Disabled Life Years: ', signif(c.dly, digits=4)))
print('-------------------')
c.dfly <- disabilityFreeLifeYearsCohort(c)
print(paste0('Cohort Disability Free Life Years: ', signif(c.dfly, digits=4)))

# Cohort disease life years
#diseaseVarLifeyear(c, 'cancre')
#diseaseVarLifeyear(c, 'diabe')
#diseaseVarLifeyear(c, 'hearte')
#diseaseVarLifeyear(c, 'lunge')
#diseaseVarLifeyear(c, 'hibpe')
#diseaseVarLifeyear(c, 'stroke')
#diseaseVarLifeyear(c, 'alzhe')
#diseaseVarLifeyear(c, 'demene')
#diseaseVarLifeyear(c, 'smoken')
#diseaseVarLifeyear(c, 'drink')
#diseaseVarLifeyear(c, 'heavy_smoker')
#diseaseVarLifeyear(c, 'problem_drinker')
```

### Smoking
```{r}
c <- cohort
# No Heart Disease
s <- smok

total_dflys_s <- s$a_disabled_all * s$m_endpop_all
sum_tot_dflys_s <- sum(total_dflys_s)
avg_dflys_s <- sum_tot_dflys_s / s$m_endpop_all[1]

avg_dflys
avg_dflys_s

### Heart Disease

```{r}

c <- cohort

total_dflys <- c$a_disabled_all * c$m_endpop_all
sum_tot_dflys <- sum(total_dflys)
avg_dflys <- sum_tot_dflys / c$m_endpop_all[1]

# No Heart Disease
h <- heart_c

total_dflys_h <- h$a_disabled_all * h$m_endpop_all
sum_tot_dflys_h <- sum(total_dflys_h)
avg_dflys_h <- sum_tot_dflys_h / h$m_endpop_all[1]

avg_dflys
avg_dflys_h

### Not Disabled
```{r}

disabilityFreeComparison <- function(c, i, int.name) {
  tot.df <- c$a_not_disabled_all * c$m_endpop_all
  sum.tot.df <- sum(tot.df)
  avg.df <- sum.tot.df / c$m_endpop_all[1]

  tot.df.i <- i$a_not_disabled_all * i$m_endpop_all
  sum.tot.df.i <- sum(tot.df.i)
  avg.df.i <- sum.tot.df.i / i$m_endpop_all[1]

  print(int.name)
  print(paste('Cohort Average Disability Free:', signif(avg.df, digits=4)))
  print(paste('Intervention Disability Free:', signif(avg.df.i, digits=4)))
  print(paste('Difference:', signif(x = (avg.df.i - avg.df), digits=4)))
}

c <- cohort
# No Heart Disease
h <- heart_c

disabilityFreeComparison(c, h, 'No Heart Disease')

disabilityFreeComparison(c, s, 'No Smoking')
```

# Defunkt
## Disability Free Life Years

@@

### Try again, this time with the correct variable (not disabled)
